---
title: "Overview of the NoaaCS package"
author: "Clément Schaff"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of the NoaaCS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, include = FALSE} 
library(NoaaCS)
library(dplyr)
library(ggplot2)
library(readr)
```

This vignette gives a brief overview of the NoasCS R package created for the purpose of visualizing NOAA earthquake data. It processes data from [NOAA database](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1) and visualizes them using `ggplot2` and `leaflet` packages. 

## Package functions

There are six exported functions available to users:

- `eq_clean_data()`
- `geom_timeline()`
- `geom_timeline_label()`
- `theme_timeline()`
- `eq_create_label()`
- `eq_map()`

## Clean data

The  function `eq_clean_data` preprocess the data. It creates a DATE column in `Date` format, transforms latitude and longitude to numeric format and trims country from LOCATION_NAME.

```{r eq_read_example, message = FALSE}
filename <- system.file("extdata/signif.txt", package = "NoaaCS")
data <- readr::read_tsv(filename)

eq_clean_data(data)
```

## Visualize earthquake timeline

Three functions use `ggplot2` to plot an earthquake timeline:

- `geom_timeline()` requires cleaned data. You should provide dates to the required aesthetics `x` , and countries to the optional aesthetics `y`. , You can set `size` and `color` according to your needs. 
- The `geom_timeline_label()` function requires additional `label` aesthetic for labeling. 
- A `theme_timeline()` is available as well. 

```{r eq_timeline_example, fig.width = 7, fig.height = 4}
data %>% eq_clean_data() %>%
     filter(COUNTRY %in% c("USA", "CHINA"), YEAR > 2000, DEATHS>0) %>%
     ggplot(aes(x = DATE,
                y = COUNTRY,
                color = DEATHS,
                size = as.numeric(EQ_PRIMARY)
     )) +
     geom_timeline() +
     geom_timeline_label(aes(label = LOCATION_NAME), n_max = 5) +
     theme_timeline() +
     labs(size = "Richter scale value", color = "# deaths")
```

## Visualize earthquakes on map

The package uses `leaflet` functions to show earthquakes on a map using `eq_map()` function. Optional annotations can be created using `eq_create_label()` function. The result is an interactive map where user can click on individual points to get details.

```{r eq_map_example, fig.width = 7, fig.height = 4}
data %>% 
  eq_clean_data() %>% 
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popup_text")
```
